---
title: "WK5 925"
author: "Zichun Liu"
date: "9/25/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nycflights13)
library(microbenchmark)
library(dplyr)
#library(tidyr)
class(flights)
```

## Why dplyr

1. Easier to read and write 
2. More efficient

```{r}
# compare efficiency
microbenchmark(
  filter(flights, month == 1, day == 1),
  subset(flights, subset = month == 1 & day == 1)
)
```


## %>% Operator

Although not required, the dplyr packages make use of the pipe operator `%>%` developed by Stefan Milton Bache in the R package `magrittr`. Although all the functions in dplyr can be used without the pipe operator, one of the great conveniences these packages provide is the ability to string multiple functions together by incorporating `%>%`.

This operator will forward a value, or the result of an expression, into the next function call/expression. For instance a function to filter data can be written as:

```{r, eval=FALSE}
filter(data, variable == numeric_value)

data %>% filter(variable == numeric_value)
```

Both functions complete the same task and the benefit of using `%>%` is not evident; however, when you desire to perform multiple functions its advantage becomes obvious. For instance, if we want to filter some data, summarize it, and then order the summarized results we would write it out as:

```{r,eval=FALSE}
#Nested Option:
arrange(summarize(filter(data, variable == numeric_value),Total = sum(variable)),desc(Total))

#or
#Multiple Object Option:
a = filter(data, variable == numeric_value)
b = summarise(a, Total = sum(variable))
c = arrange(b, desc(Total))

#or
#%>% Option:
data %>%
  filter(variable == “value”) %>%
  summarise(Total = sum(variable)) %>%
  arrange(desc(Total))
```

As your function tasks get longer the `%>%` operator becomes more efficient and makes your code more legible. In addition, although not covered in this tutorial, the %>% operator allows you to flow from data manipulation tasks straight into vizualization functions (via ggplot and ggvis) and also into many analytic functions.


## dplyr functions

```{r,eval=FALSE}
select() # select columns # Reduce dataframe size to only desired variables for current task

filter() # select rows # Reduce rows/observations with matching conditions

group_by() # Group data by categorical variables

summarise() #Perform summary statistics on variables

arrange() # Order variable values

XXX_join() # Join two datasets together

mutate() # Creates new variables
```

```{r}
#select
flights %>%
  select(month, day, carrier, distance)
```
```{r}
# filter
flights %>%
  select(month, day, carrier, distance) %>%
  filter(month == 1, day == 1)
```

```{r}
# summarise and group_by
flights %>%
  select(month, day, carrier, distance) %>%
  filter(month == 1, day == 1) %>%
  group_by(carrier) %>%
  summarise(meandistance = mean(distance))
```


```{r}
#arrange
flights %>%
  select(month, day, carrier, distance) %>%
  filter(month == 1, day == 1) %>%
  group_by(carrier) %>%
  summarise(meandistance = mean(distance)) %>%
  arrange(desc(meandistance))
```
```{r}
# mutate()
flights %>%
  select(month, day, carrier, distance) %>%
  filter(month == 1, day == 1) %>%
  group_by(carrier) %>%
  summarise(meandistance = mean(distance), N = n()) %>%
  arrange(desc(meandistance)) %>%
  mutate(totaldistance = meandistance * N)
```


```{r}
(flight_1 = flights %>%
  select(month, day, carrier, distance) %>%
  filter(month == 1, day == 1) %>%
  group_by(carrier) %>%
  summarise(meandistance = mean(distance), N = n()) %>%
  arrange(desc(meandistance)) %>%
  mutate(totaldistance = meandistance * N))
```

```{r}
(flight_2 = flights %>%
  select(month, day, carrier, air_time) %>%
  filter(month == 1, day == 1) %>%
  group_by(carrier) %>%
  summarise(meantime = mean(air_time), N1 = n()) %>%
  arrange(meantime) %>% na.omit())
```
```{r}
inner_join(flight_1, flight_2, by = "carrier")
left_join(flight_1, flight_2, by = "carrier")
full_join(flight_1, flight_2, by = "carrier")
semi_join(flight_1, flight_2, by = "carrier")
anti_join(flight_1, flight_2, by = "carrier")
```

```{r}
library(ggplot2)
inner_join(flight_1, flight_2, by = "carrier") %>%
  ggplot() + 
  geom_bar(aes(x = carrier, y = meandistance, fill = N), stat = "identity")


inner_join(flight_1, flight_2, by = "carrier") %>%
  ggplot() + 
  geom_point(aes(x = meandistance, y = meantime, col = carrier)) +
  geom_smooth(aes(x = meandistance, y = meantime), method = "lm", lwd = 0.5)
```


## End

```{r, include=FALSE, eval=FALSE}
## tidyr functions
gather() # takes multiple columns, and gathers them into key-value pairs: it makes “wide” data longer
spread() # takes two columns (key & value) and spreads in to multiple columns, it makes “long” data wider
separate() # splits a single column into multiple columns
unite() # combines multiple columns into a single column
```
