---
title: "dplyr"
author: ""
date: "September 22, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nycflights13)
library(tidyverse)
```

## Data Manipulation

Pick observations by their values (filter()).  

Reorder the rows (arrange()).  

Pick variables by their names (select()).  

Create new variables with functions of existing variables (mutate()).

Collapse many values down to a single summary (summarise()).

## All verbs work similarly:  

The first argument is a data frame.  

The subsequent arguments describe what to do with the data frame, using the variable names (without quotes).  

The result is a new data frame.  

## OK, but what is a tibble?

[Look here for everything you want to know](http://bit.ly/2ffxvfM)  


Moving on with dplyr ...  

## filter()

filter() allows you to subset observations based on their values.

```{r}
filter(flights, month == 1, day == 1)

```


## R either prints out the results, or saves them to a variable.

If you want to do both, you can wrap the assignment in parentheses:

```{r}
(jan1 <- filter(flights, month == 1, day == 1))

```

## NA -- key knowledge for statisticians

```{r}
NA > 5

10 == NA

NA + 10

NA / 2

## -------------------

NA == NA

```

## arrange()

arrange() works similarly to filter()   
except that instead of selecting rows,   
it changes their order.

```{r}
arrange(flights, year, month, day)
```


## desc()

Use desc() to re-order by a column in descending order:

```{r}
arrange(flights, desc(arr_delay))
```


